---
title: "Sex Discrimination in Employment"
author: "Kaylhan Garcia"
date: "2025-01-28"
output: word_document
---

# **Sex Discrimination in Employment**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load appropriate libraries and data

```{r}
library(tidyverse)
library(Sleuth3)
library(psych)
data(case0102)

help(case0102)
```

## Let's summarize the data by gender

```{r}
# Using a pipe describe Salaries by sex
case0102 %>%
  group_by(Sex) %>%
  summarise(n=n(), Average=mean(Salary), SD=sd(Salary),
            Median=median(Salary), IQR=IQR(Salary),
            Min=min(Salary), Max=max(Salary),
            Q1=quantile(Salary,0.25), Q3=quantile(Salary, 0.75)) %>%
  knitr::kable(digits = 3)

```

## Now let's see the data across both genders

```{r}
describe(case0102$Salary , skew= FALSE, ranges = TRUE, IQR = TRUE, omit = TRUE)

```

##Mutating the data for "Male" to appear first

```{r}
case0102 <- case0102 %>%
  mutate(Sex = factor(Sex, levels=c('Male', 'Female')))
levels(case0102$Sex)
```

## Begin visualizations

```{r}

ggplot(data = case0102, mapping = aes(y = Salary, x = Sex)) +
  geom_boxplot(fill="blue")+
  ylab("Salary")+
  xlab("Gender")
```

```{r}
ggplot(data = case0102, mapping = aes(x = Salary, fill = Sex)) +
  geom_density(alpha = 0.3)+
  xlab("Salary")+
  theme_bw()
```

This is a **density plot** comparing salaries between **males (pink) and females (blue).**

Instead of showing exact counts, it shows the **shape of the salary distribution**—where salaries are most concentrated.

-   Female salaries tend to cluster around **\$5000**, while male salaries peak around **\$6000.**

## Hypotheses Testing

### Let's use a T-Test

```{r}
t.test(Salary ~ Sex, var.equal = TRUE, data = case0102)
```

## Inference

We are 95% confident that the true difference in mean salaries between males and females lies between \$560 and \$1,076

**Male mean:** \$5956.88

**Female mean:** \$5138.85

**Difference:** \$818.03

H0: There is no difference in the mean salaries between males and females. #Ha: There is a difference in the mean salaries between males and females.

The test shows a statistically significant difference in mean salaries between males and females


